\section{Application Protocol Middleware}
\label{sec:apps}

In this section, we discuss the four different classes of applications that can be built on the Latitude platform. Each
of these applications is supported by a suite of software modules, smart contract add-ons and SDKs (mobile and desktop)
that utilize the core functionality offered by Latitude and provision it in unique ways to suit the specific class of
applications. In the first few versions, these applications are supported by a middleware layer as shown in Figure
\ref{fig:lat-arch}. As Latitude grows, we shall be able to scale by moving to a side-chain architecture, details of
which are discussed in Appendix \ref{app:latchain}.

The Latitude platform is extensible in the sense that its not limited to these four classes of application layers. For
example, its possible to add additional applications such as for the shipping or the airline industry as verticals on
the platform. Next we discuss each of these applications and how they can be supported on the platform.

\subsection{Ride-share application layer}

Ride-sharing applications are the pinnacle of the applications in the data sharing economy. They empower the user to
choose from a multi-modal ways of finding or providing rides. Latitude can provide the decentralized infrastructure to
store, share and build decentralized applications for their corresponding counterparts.

The ride-sharing market alone is large enough to power the Latitude blockchain as its primary use-case. The ride-sharing
market is around 17 Billion dollars with around 60 Million users. This does not include the multi-modal ride sharing
segment which is growing rapidly. This segment includes bike, scooter and such modes of transport. The ARPU is around 
293 dollars which is high enough to create user incentivized models for data sharing. The data generated by these users
currently sits in silos in the respective ride sharing apps, which can be unlocked and put to good use through such
mechanisms. 

Latitude shall provide a mobile SDK and a blockchain API specifically to suit ride-sharing apps for sharing data. This
would allow the creation of decentralized ride sharing apps that can provide users with incentives to share data,
subsidize rides and provide a better deal for ride providers. They can also allow regulators to enforce the use of
roads, lanes, parking and other structures in accordance with city or neighborhood ordinances. Latitude's SDK shall also
provide a multi-modal ride API that such apps can use to request rides on the platform. 

An app that integrates with the ridesharing SDK becomes a trusted data provider. Further verification of the letigmacy
of the data source can be performed by integrating with Trusted Execution Environments (TEEs) available on most iOS and
Android phones today. Trusted data sources provide an increased amount of trust in as given in the proof of X system for
Latitude.

\noindent
{\em User rewards:} The design of the ride-share protocol stack shall include token rewards for contributing data. The
reward mechanism shall be provided using a vetted smart contract system. The spatial datastore shall be used for hosting
the data. The protocol stack shall also host applications that use the data in accordance with the data contract between the
data providers and the app developers.

As an preferred third-party integration, Latitude shall integrate with SherpaShare's platform for drivers. This would
allow greater sharing of the geo-spatial and driving data collected through SherpaShare's driver app. This will be one
of the early launches for the beta version of the Latitude platform.

In order to scale, the Ride-sharing protocol can use a side-chain design where governance rules can be baked into the
side-chain. This is further discussed in Appendix \ref{app:sidechain}.

Some of the participants on the ride-sharing protocol include:

\begin{itemize}
    \item Multi-modal ride sharing data sources: Bike and scooter apps such as LimeBike, Mobike, OFO, BlueGoGo, Youon,
        MingBikes, etc.
    \item Traditional ride-sharing data silos: Examples include Uber, Lyft, Gett, Uno, Curb, Sitbaq, etc.
    \item Public transit data sources: Bus, train and other transport data.
    \item Ride-sharing dapps: Apps can utilize such data for incentivizing users, creating ride sharing applications or
        other creative uses.
\end{itemize}

%Overview of ridesharing applications. Big industry leaders. Statistics on rides, miles, revenue, ARPU.
%
%Multi-modal ride data:
% - Bike, scooter ride apps.
%Disruptive to Bikesharing:
% - Mobike, OFO, BlueGoGo, Youon, Mingbikes
% - Hellobike, YooBike, CCbike, Zagster, LimeBike
% - Citi Bike, Capital Bikeshare, Divvy, Hubway, Docomo Bike
%Share, Relay Bikes
% - Public transit data
%Use SDK/app to share data. Similar incentive model.
%Monetization:
% - Single multi-modal ride API (SherpaShare or 3rd party).
%
% Ridesharing platform:
%
% Platform:
%Latitude Ridesharing SDK.
%Ridesharing sidechain on the Latitude blockchain.
%Riders:
%Contributing ridesharing data:
%Uber, Lyft, Gett, Juno, Curb, Sitbaq
%Drivers:
% - Use SherpaShare app or 3rd-party app with SDK.
% - Proof of ride:
%Using driver-side app and/or client app.
%Other players:
% - City, Law enforcement, analytics use cases.

\subsection{Telematics Application Layer}

Data is the most important asset for Telematics applications. Thus, sharing of data, computations and algorithms can be
beneficial to companies, users and the ecosystem in general. In fact, by using the Latitude blockchain it becomes
possible for regulators to enforce laws better while maintaining data privacy. This can result in lower crime rates and
incidents.

The important of data sharing here can be seen from the existence of {\em Telematics data exchanges} that are
centralized cloud-based exchanges which act as data brokers among users and insurance companies. As an example, consider
the Verisk Data Exchange (verisk.com) which allows the exchange of driver data to insurance companies. They share data
for all kinds of connected vehicles for underwriting, rating, and claims handling through Usage-based Insurance (UBI)
programs.  The exchange stores and processes Telematics data of all types, volumes, and velocity from connected cars,
after-market hardware, or mobile solutions. A second example is the Octo Telematics company which also operates an
exchange that is used by over 100 insurance companies, has about 186 billion miles of driving data and gets 11 billion
new data points from 5.4 million connected cars and sensors every day.

The problem with such centralized solutions is that there is a central entity that extracts the fees and has direct
control over the data. Such centralization can result in security and privacy loss and lack of trust from all parties
involved. Latitude is directly poised to disrupt this market by providing a blockchain based platform for exchange of
data, computation, algorithms and resulted information. Using Latitude smart contracts it becomes possible to enforce
proper security, privacy and sharing of data in accordance with agreed upon incentives. This implies cleaner and better
sharing methods which can be a win-win for everyone in the industry including the regulators.

Latitude shall provide a Telematics SDK and blockchain library module to enable such applications. This would include a
suite of pre-existing smart contracts, datastore provisions and other software modules necessary to build decentralized
data exchanges which can eliminate the expensive data brokers. The SDK shall also include APIs for Usage-based Insurance
companies to act as data providers given appropriate involvement, permission and incentives for the users. Users can
also be incentivized to provide data directly to the platform by getting token rewards. Such a model also allows for
open experimentation with driver behavior algorithms such as DriverScore. It also becomes possible for drivers and users
to "carry" their score or reputation from one platform to another, thus preventing lock-in scenarios.

%---Telematics data sharing:
%Data sharing can result in better algorithms, better understanding among drivers.
%Telematics data can save lives:
%Allows government regulations to be enforced. Data privacy can be enforced.
%Driver’s own their driving history (score) and can choose to carry it.
%
%
%---UBI methods can be improved with data sharing and incentives.
%Example:
%Octo telematics
%20 million miles, 186 billion miles of driving data
%100 insurance companies (data is not shared).
%11 billion new data points daily from 5.4 million connected cars and sensors.
%Disruptive opportunity for data sharing.
%
%
%
%---Telematics data exchanges exist today !
%But centralized, unclear policies, enforcement.
% Lack of control.
% Example: Verisk Data Exchange (verisk.com)
% The first-of-its-kind  exchange draws driving data from all kinds of connected vehicles for underwriting, rating, and
% claims handling through UBI programs. 
% The exchange is scalable to normalize, process, and store telematics data of all types, volumes, and velocity from
% connected cars, after-market hardware, or mobile solutions.
% Finally, one exchange solves the many-to-many challenge by connecting auto-makers and telematics service providers
% (TSPs) to multiple insurers
%
%
% ---Insurance Vertical
% Latitude Driver SDK:
% Allows trusted capture of driving data.
% Platform for running DriverScore / Behavior algorithms on phone or in the cloud (or a combination). 
% Eg: Geico:
% Can run 3rd-party algorithms such as Geico’s driver score directly on the phone.
% Includes API for interacting with Geico’s app.
% Users mine currency by:
% Contributing driver data, driver scores.
% Users own the data, have full control.
%
% Insurance providers (Geico, Progressive, Farmers, etc)
% Use platform for computing metrics.
% Contribute data/and or crypto-currency incentivized.
% Always user permissioned.
% Other players:
% City, Law enforcement, analytics use cases.
% Car history use cases (better version of Carfax).

\subsection{Mapping and Location Application Layer}

Mapping and Location-based services and analytics is another big segment that Latitude can address through
decentralization. Mapping and real-time location/speed data is a huge market that apps like Waze are able to access.
However, the user incentives in Waze are limited to the Waze platform and cannot be carried over. Finally, the data that
users contribute also stays isolated to the Waze platform and cannot be used by the City or the National Transportation
Safety Board (NTSB) for altruistic purposes. To get a sense of the market size, today Waze has about 70 million users
with 500K volunteers who provide real-time data. A Waze user on average spends about 480 minutes per month on the
platform.

The Latitude blockchain can change all of data isolation and platform lock-in by providing strong incentives to the user
by rewarding them with crypto-currencies. The smart contract system can also be used to enforce proper sharing of data
with the right participants. The built-in primitives in Latitude for Byzantine behavior can be used to ensure honest
operation at all times.

Location-based services can be similarly decentralized on Latitude. Using the mobile Latitude SDK, it becomes possible
for users to directly contribute mapping and location data to the platform. This allows Latitude to construct the
various cryptographic proofs as discussed in Section \ref{sec:crypto}. Proof of location technology alone has the
potential to disrupt the "Check-ins" industry. The Facebook platform alone gets about 50 million check-ins per year. The
other major players here include Foursquare, Yelp and Google maps. The sharing of check-ins data on the Latitude
platform can enable new applications not previously possible due to the platform lock-ins.

Here are some of the features present in this layer:

\begin{itemize}
    \item SDK for integrating with mobile apps that provide mapping/location data and related proofs. These SDKs
        integrate with smart contracts built as part of the protocol layer on Latitude to support mapping/location.
    \item SDK contains APIs for various mapping/location use cases, such as providing landmark data, providing location
        data and proofs, traffic incident reports to name some.
    \item Protocol layer for building apps: The mapping/location protocol layer includes APIs for dapp developers to
        build apps on this ecosystem that can utilize the mapping/location data. Apps can include real-time traffic,
        incident monitoring or prediction. Utilizing location proof for access control or other applications such as
        check-ins. Building a decentralized Waze is a canonical exmaple here.
    \item Datastore API: The protocol layer includes a datastore API to query and index mapping/location data for
        efficient retrieval.
\end{itemize}

%Waze example:
% - 70 Mil users, 500k volunteers share data. user spends 480 min/month.
%
%-- Upcoming mapping applications include:
% High-resolution data for self-driving technology.
% Road data, needs consensus and trust.
% Proof of location:
% Location enabled maps. 
% Foursquare checkins. Facebook checkins.
% Location-based access control.
% 3 BILLION location requests daily on Android
% 250M location-enabled service users.
%
%--Latitude Mapping + Location SDK
%Mobile SDK for trusted mapping + location.
%SherpaShare app + 3rd party apps.
%Users mine by:
%Contributing road, traffic, street, mapping data.
%Contributing user location
%Proof of location concept can replace “Check-Ins”
%Facebook has 50 million check-ins per year.
%City, law enforcement, analytics use cases.
%User has full privacy control.
%
%-- Applications:
%Waze like apps:
%By Latitude Mapping/Location Sidechain.
%Waze, OsmAnd, Maps.Me, HEREWeGo, 2GIS
%Yandex.Maps, Locus Maps.
% Navmil, NavIt, MapQuest.
% Crowd-sourced maps.
% Authenticated location applications:
% Access control or privileges by virtue of location.
%
\subsection{Smart city and Government Application Layer}

Smart cities refer to a large set of applications that can improve the life of a citizen by providing better amenities.
Examples include real-time data for parking, data-driven urban planning, city-wide zoning research,
environmental location sensors, disaster management, transport/transit data (real-time and historical), smart transport
(shuttle/bikes), traffic light management to name a few.

Latitude blockchain can host such smart city applications and collect data from their citizens. Users are incentivized
to provide data using token economics and the regulators or city officials can "purchase" data or results using smart
contracts. It might be possible to share aggregated data with careful focus on privacy/anonymity to State level or
Federal government for various regulatory reasons. 

\begin{itemize}
    \item City can use traffic data for to improve planning and decisions for roads, HOV lanes, traffic lights and
        parking. The algorithms that process this data will be controlled using smart contracts that run on behalf of
        the city on the network using the Latitude protocol layer for smart city. The cost of running the smart contract
        gets payable using the LAT token.
    \item The city can also use the platform to institute regulations such as reporting of accidents or traffic
        incidents. In this regard, citizen safety is an application that can be built on this layer. Safety includes
        proof of location in real-time integrated with Police or Fire departments.
    \item Bike ride-sharing: Bike ride-sharing has seen a recent boom, but are also facing regulatory backlash. Using
        Latitude, bike ride-share providers can share the current locations of where bikes are parked. City officials
        can designate "virtual" no-bike zones which can be enforced on the chain using contracts and proofs. 
\end{itemize}

%One of the final set of applications.
%Smart city applications:
% - Parking, data-driven urban planning, waste management, water software and analytics, environmental location sensors,
%   disaster management, transport/transit data (real-time and historical), smart transport (shuttle/bike),
%   connectivity citywide, grid/energy, traffic light management, security and survelliance.
%
%  Users mine by contributing data.
%  Govt requests access using crypto.
%  Use cases:
%  Census, traffic, analytics, city-wide zoning research.
%  Real-time understanding of citizen data.
%  Compliance:
%  Insurance, Ridesharing companies regulatory compliance enforced using smart contracts on the Latitude blockchain.

\section{Implementation Roadmap}
\label{sec:roadmap}

In this Section, we discuss an implementation roadmap for Latitude for the next year. For the initial development, we
shall use an existing blockchain for the base layer, such as Ethereum, EOS or Neo. The design shall include an
abstraction layer so that later we can fit in our own Latitude Blockchain to support all operations. The following is
the roadmap for development. Note that this development happens in parallel with customer acquisition, platform
promotion and other business development activities:

\begin{enumerate}
    \item Ride-sharing protocol layer: Build a Latitude SDK and blockchain protocol layer for ride-sharing data. Build a
        simple smart contract substrate to support ride-sharing data collection.
    \item Spatial datastore: Build ability to store and retrieve location, mapping, telematics and ride-sharing data.
    \item Latitude smart contract layer: Using WASM, build custom libraries for Latitude smart contracts. Build APIs for
        Ride-sharing or mapping/location as the first application layer.
    \item Proof of X: Build smart contracts for proof of location, mapping and others.
    \item SDK Developer launch: Launch hackthons to show case proof of location, ride-share, mapping and other forms of
        data sharing.
    \item Regulatory smart contracts: Demonstrate implementation of city/state regulations on transportation data.
    \item Testnet launch: Launch testnet with above capabilities. Grow developer community with 1000 smart contracts and
        dapps written on the platform.
    \item Governance Layer: Create a list of governance operations. Build smart contracts that allow governance
        decisions to be made. Create an initial list of governance nodes and membership parameters. Launch on testnet.
    \item Latitude Blockchain: Build the core blockchain, governance layer and mining code. Update testnet with core
        blockchain capabilities. Iterate, bugfix and get feedback from community.
    \item Launch mainnet: Launch mainnet with own Latitude blockchain, SDKs for each application layer, 1000+ smart
        contract library, dapps and other developer tools.
\end{enumerate}
